---
title: "POLS 3000 Overview of Major Commands"
output: html_document
---

In this handout, we use a data frame `df`: 

```{r}
# Create a toy data frame `df`:
df <- data.frame(var1 = c("landline","Internet1","Internet2", "cell", "landline","Internet1","Internet2", "cell"), var2 = c(40,50,60,70,40,50,60,70), var3 = c(60,50,40,60,60,80,40,78), var4 = c("yes","yes","yes","yes","no","no","no","no"))

# We have created:
df # normally, we do not want to print an entire data frame, but this one is very small

# Obtain a summary:
summary(df)

# Which variables are factor ones?
class(df$var1)
class(df$var2)
class(df$var3)
class(df$var4)
```

# `table()` function

* Use to tabulate two categorical variables

* `table()` function takes in the names of the variables that you want to tabulate 

* Example 1: `table(df$var1, df$var2)`, i.e., arguments are the names of the variables that you want to tabulate: 

* Using `table()` with continuous variables is not as informative, instead use `tapply()`

```{r}
table(df$var1, df$var2)

## Make the table more informative: 
addmargins(table(InformativeLabel1 = df$var1, InformativeLabel2 = df$var2))
```

* Example 2: `table(df$var1)` will simply return the frequencies of observations (i.e., rows) in each category of `var1`
```{r}
table(df$var1)
```


# `tapply()` function

* Use to apply a function to a continuous variable across certain categories

* Example 1: `tapply(df$var3, df$var1, mean)`. This code reads: calculate the average of var3 for each category of var1. 

```{r}
tapply(df$var3, df$var1, mean)
```

* Example 2: `tapply(df$var3, df$var4, quantile)`. This code reads: calculate the quantiles of var3 for each category of var4. 

```{r}
## Make the table more informative: 
tapply(df$var3, df$var4, quantile)
```


# `subset()` function

* Use to create smaller data frames that satisfy certain conditions

* Takes in three arguments: 1) the name of the old data frame from which we will choose certain rows/columns; 2) the logical statement that specifies which rows to grab; 3) the logical statement that specifies which columns to grab

* Example 1: `subset(df, subset = (var4=="no" | var1 == "Internet1"), select=c(var2, var3))`. This code reads: 1) take the old data frame `df`; 2) only grab the rows from `df` that either equal to "no" (var4) OR equal to "Internet" (var1); 3) only grab two columns: var2, var3

```{r}
df.new <- subset(df, subset = (var4=="no" | var1 == "Internet1"), select=c(var2, var3))

summary(df.new)
```


* Example 2 (grabbing only rows): `subset(df, subset = (var4=="no" | var1 == "Internet"))`. This code reads: 1) take the old data frame `df`; 2) only grab the rows from `df` that either equal to "no" (var4) OR equal to "Internet" (var1)
```{r}
df.new1 <- subset(df, subset = (var4=="no" | var1 == "Internet1"))

summary(df.new1)
```


* Example 3 (grabbing only columns): `subset(df, select=c(var2, var3))`. This code reads: 1) take the old data frame `df`; 2) only grab two columns: var2, var3
```{r}
df.new2 <- subset(df, select=c(var2, var3))

summary(df.new2)
```



# `ifelse()` function

* Use to recode variables

* Takes in three arguments: 1) the logical statement; 2) the value to be assigned if the logical statement is TRUE; 3) the value to be assigned if the logical statement is FALSE

* Example 1: `ifelse((var4=="no" | var1 == "Internet1"), 1, 0)`. This code reads: 1) evaluate whether an observation either has "no" in var4 OR has "Internet1" in var1; 2)  if TRUE, assign 1; 3) if FALSE, assign 0

```{r}
df$var5 <- ifelse((df$var4=="no" | df$var1 == "Internet1"), 1, 0)

df 
```


* Example 2: `ifelse((var4=="no" | var1 == "Internet1"), "Good", "Bad")`. This code reads: 1) evaluate whether an observation either has "no" in var4 OR has "Internet1" in var1; 2)  if TRUE, assign "Good"; 3) if FALSE, assign "Bad""

```{r}
df$var6 <- ifelse((df$var4=="no" | df$var1 == "Internet1"), "Good", "Bad")

df
```

* Example 3: we will create three as opposed to two categories. For instance, consider: `ifelse((df$var4=="no"), "Good", ifelse((df$var1 == "Internet1"), "Bad", NA)`. This code reads:  evaluate whether an observation  has "no" in var4, if TRUE, assign "Good", if FALSE evaluate whether an observation has "Internet1" in var1, if TRUE, assign "Bad", otherwise assign NA. 

```{r}
df$var6 <- ifelse((df$var4=="no"), "Good", 
                  ifelse((df$var1 == "Internet1"), "Bad", NA))

df
```
