---
title: "POLS 3000: Problem Set 1"
author: "FirstName LastName"
output: html_document
date: "September 8, 2017"
---

General guidelines: make sure to use comments to explain what you intended the code 
to do. Use "Insert > R" to insert your R code. Use "Run > Run Selected Lines" to 
check if your code works. Use "Knit" to compile the HTML output file.

## Exercise 1.5.1

Surveys are frequently used to measure political behavior such as
voter turnout, but some researchers are concerned about the accuracy
of self-reports.  In particular, they worry about possible *social
desirability bias* where in post-election surveys, respondents who did
not vote in an election lie about not having voted because they may
feel that they should have voted.  Is such a bias present in the
American National Election Studies (ANES)?  The ANES is a nation-wide
survey that has been conducted for every election since 1948.  The
ANES conducts face-to-face interviews with a nationally representative
sample of adults. Table  1.3 displays the names and descriptions
of variables in the `turnout.csv` data file.

 <!-- Make sure to read Table 1.3, also shown below -->
 <!-- If you want to uncomment the table, go to Code > Comment/Uncomment Lines -->
 
 Name                 Description
 -------------------- ----------------------------------------------------------
 `year`               Election year
 `VEP`                Voting Eligible Population (in thousands)
 `VAP`                Voting Age Population (in thousands)
 `total`              Total ballots cast for highest office (in thousands)
 `felons`             Total ineligible felons (in thousands)
 `noncitizens`        Total non-citizens (in thousands)
 `overseas`           Total eligible overseas voters (in thousands)
 `osvoters`           Total ballots counted by overseas voters (in thousands)


### Question 1.5.1.1

a) Load the set `turnout.csv` data into R.  

b) Check the dimensions of the data.

c) Obtain a summary of the data.  

d) How many observations are there? What is the range of years covered in this data set? 

### Answer 1.5.1.1

```{r}

```



### Question 1.5.1.2

  A turnout rate is a percentage of voters who voted (nominator) out of all potential 
  voters (denominator) multiplied by 100. For instance, if 2 out of 10 potential voters voted, 
  then the turnout rate is: $\displaystyle\frac{2}{10}\times100=20\%$. 

a) Calculate the turnout rate as a proportion of the entire voting age population. 
  Note that for this data set, we must add the total number of eligible overseas 
  voters since the *VAP* variable does not include these individuals in the count. 

b) Next, calculate the turnout rate using the voting eligible population or VEP.  

c) Extra credit (1%): add labels for each turnout rate, using the year of election.

d) Produce two plots such that both turnout rates are displayed over years.

e) Extra credit (1%): search online how to limit the y-axis such that plots appear 
  side by side and limit the y-axis to a range that you consider appropriate. Make
  sure the y-axis is the same in both plots, to make the comparison easy.

f) Summarize what difference you observe in 1-2 sentences.
  
### Answer 1.5.1.2

```{r}

```



### Question 1.5.1.3

Recall that American National Election Studies (*ANES* variable) records 
the turnout rate of American voters by conducting face-to-face interviews 
with a nationally representative sample of adults. This raises the concern 
about possible *social desirability bias*.
 
a. Compute the difference between VAP and ANES estimates of turnout rate.  
    How big is the difference on average?  What is the range of the difference?  
  
b. Conduct the same comparison for the VEP and ANES estimates of voter turnout.  
    How big is the difference on average?  What is the range of the difference?  

c. Summarize your findings in 5-6 sentences.

### Answer 1.5.1.3

```{r}

```




## Exercise 2.8.3


One longstanding debate in the study of international relations
concerns the question of whether individual political leaders can make
a difference.  Some emphasize that leaders with different ideologies
and personalities can significantly affect the course of a nation.
Others argue that political leaders are severely constrained by
historical and institutional forces.  Did individuals like Hitler,
Mao, Roosevelt, and Churchill make a big difference?  The difficulty
of empirically testing these arguments stems from the fact that the
change of leadership is not random and there are many confounding
factors to be adjusted for.

In this exercise, we consider a *natural experiment* in which the
success or failure of assassination attempts is assumed to be
essentially random.

This exercise is based on:
Jones, Benjamin F, and Benjamin A Olken. 2009. “[Hit or Miss? 
 The Effect of Assassinations on Institutions and 
 War.](http://dx.doi.org/10.1257/mac.1.2.55)” 
 *American Economic Journal: Macroeconomics* 1(2): 55–87. 

Each observation of the CSV data set
`leaders.csv` contains information about an assassination
attempt.  The variables are:

 Name                 Description
 -------------------- ----------------------------------------------------------
 `country`            The name of the country
 `year`               Year of assassination
 `leadername`         Name of leader who was targeted
 `age`                Age of the targeted leader
 `politybefore`       Average polity score during the 3 year period prior to the attempt
 `polityafter`        Average polity score during the 3 year period after the attempt
 `civilwarbefore`     1 if country is in civil war during the 3 year period prior to the attempt, or 0 
 `civilwarafter`      1 if country is in civil war during the 3 year period after the attempt, or 0
 `interwarbefore`     1 if country is in international war during the 3 year period prior to the attempt, or 0
 `interwarafter`      1 if country is in international war during the 3 year period after the attempt, or 0
 `result`             Result of the assassination attempt, one of 10 categories described below

The `polity` variable represents the so-called *polity score*
from the Polity Project.  The Polity Project systematically documents
and quantifies the regime types of all countries in the world from
1800.  The polity score is a 21-point scale ranging from -10
(hereditary monarchy) to 10 (consolidated democracy).  

The `result` variable is a 10 category factor variable describing
the result of each assassination attempt.

### Question 2.8.3.1

a. Load the data set `leaders.csv`.

b. How many assassination attempts are recorded in the data?  

c. How many countries experience at least one leader assassination attempt?
   (The `unique()` function, which returns a set of unique values from the input 
   vector, may be useful here).  

d. What is the average number of assassination attempts per year? 
   (The `table()` function, which returns  may be useful here).
   
e. Summarize your findings in 3 sentences.   

### Answer 2.8.3.1

```{r}

```



### Question 2.8.3.2

a. What are the levels of the variable `result`? Save the levels as a new array object.

b. Using indexing of the newly created array object, create a new binary variable named 
   `success` that is equal to 1 if a leader dies from the attack and to 0 if the leader survives.  
   Using `ifelse()` is useful here, make sure to reread the descrition in the textbook.
   
c. Store this new variable as part of the original data frame.  
   
d. What is the overall success rate of leader assassination?

e. Summarize your finding from section d. as in one sentences. 


### Answer 2.8.3.2

```{r}

```


### Question  2.8.3.x (not from the textbook)

a. Subset the `leaders` data set to create four subsets that contain only: 
   - the attempts that followed either an interstate war or a civil war,
   - the attempts that followed an interstate war, 
   - the attempts that followed a civil war, 
   - the attempts that followed no wars.
   
b. What is the average success rate for these four subsets of cases?

c. Compare the success rates in the entire dataset and the four subsets of cases.

### Answer 2.8.3.x
```{r}


```
